---
title: "TITLE PROJECT 2"
author: "Team 3; Group 1"
date: "02/09/2021"
output:
  html_document:
    df_print: paged
    number_sections: yes
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

***

# Abstract 


# Introduction

 
# Background 


# Descriptive analysis 


# Inferential analysis 

##depiction of the models we use in statistical terms

# Sensitivity analysis 

#Causal Inference

# Discussion 

<span style='color:blue'> 
Conclude your analysis in this section. You can touch on the following topics. 
</span> 

- A brief recap of this project. 
- Findings in the inferential analysis interpreted in the context of Project STAR. 
- Suggestions for future research and/or policy making given your findings. 
- Caveats of the current analysis.

# Acknowledgement {-}

<span style='color:blue'>
By default, it is assumed that you have discussed this project with your teammates and instructors. List any other people that you have discussed this project with. 
</span>

# Reference {-}


# Session info {-}

<span style='color:blue'>
Report information of your `R` session for reproducibility. 
</span> 

```{r}
sessionInfo()
```